# Install packages
```{r}
install.packages('ggthemes', dependencies = TRUE) 
install.packages('ggplot2', dependencies = TRUE) 
install.packages('knitr', dependencies = TRUE) 
library(ggplot2)
library(ggthemes)
library(knitr)
```
# Data of birth plot
## First create with breaks
```{r}
ggplot(aes(x = dob_day), data = pf) + 
  geom_histogram() + 
  scale_x_discrete(breaks = 1:31)
```
## Date of birth facebook data, splitting into 2 different months
```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram() + 
  scale_x_discrete(breaks = 1:31) + 
  facet_wrap(~dob_month)
```
### Change directory - get Facebook data
```{r}
setwd("~/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson3")
pf <- read.delim('pseudo_facebook.tsv')
```

# Plot Friend Count Facebook data
```{r}
names(pf)
qplot(x = friend_count, data = pf)
```

## answer in ggplot form
```{r}
ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram()
```
## Plot friend count with different binwidth/setting x scale with breaks (split gender into different plots) 
```{r}

qplot(x = friend_count, data = pf, binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
facet_wrap(~gender)
```
## alternative answer using facet_grid
```{r}
qplot(x = friend_count, data = pf) + 
  facet_grid(gender ~ .) 
```
## alternative answer in ggplot2
```{r}
ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

## NA values - http://www.statmethods.net/input/missingdata.html
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
## another way to deal with Na values - caution: more dangerous as it will omit
# NA values in other variables, not only gender
```{r}
qplot(x = friend_count, data = na.omit(pf), binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
## in ggplot2
```{r}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```


# Statistics 'by' Gender Solution
```{r}
by(pf$friend_count, pf$gender, summary)
```

The parameter color determines the color outline of objects in a plot. 
The parameter fill determines the color of the area inside objects in a plot. 
You might notice how the color black and the hex code color of #099DD9 (a shade of blue)
are wrapped inside of I(). The I() functions stand for 'as is' and tells qplot 
to use them as colors. 

# ggplot2 docs on themes: http://docs.ggplot2.org/0.9.2.1/theme.html
```{r}

ggplot(aes(x = tenure), data = pf) + 
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

# Tenure
```{r}
qplot(x = tenure/365, data = pf, binwidth = 1, color = I('black'), fill = I('#099DD9'))
```
# Improved plot
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq (1, 7, 1), lim = c(0,7)) # makes breaks go from 0-7
```
## ggplot2 equivalent
```{r}
ggplot(aes(x = tenure/365), data = pf) + 
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

# Adding labels to plots
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25, color = I('black'), fill = I('#F79420'),
  xlab = "Number of years using Facebook",
  ylab = "Number of users in sample") +  
  scale_x_continuous(breaks = seq (1, 7, 1), lim = c(0,7)) # makes breaks go from 0-7
```
# Add nice labels to the plot
## ggplot2 version
```{r}
ggplot(aes(x = tenure / 365), data = pf) + 
  geom_histogram(color = 'black', fill = '#F79420') + 
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) + 
  xlab('Number of years using Facebook') + 
  ylab('Number of users in sample')
```


# Exercise to create Histogram of ages from pf dataset
```{r}
names(pf)
ggplot(aes(x = age), data = pf) +
  geom_histogram( color = 'black', fill = 'yellow') +
  scale_x_continuous(breaks = seq(1, 120, 4), limits = c(10,120)) +
  xlab("Age in years") +
  ylab("Number of users in sample")
```
## Instructor's plot
```{r}
ggplot(aes(x = age), data = pf) + 
  geom_histogram(binwidth = 1, fill = '#5760AB') + 
  scale_x_discrete(breaks = seq(0, 113, 5))
```

# Exercise to use scales and create multiple plots on one page. Use log10 & sqrt

Tips about transforming data with log base 10 and log base 2: 
[R-stats blog](http://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/)
```
signedlog10 = function(x) {
ifelse(abs(x) <= 1, 0, sign(x)*log10(abs(x)))
}
```
```{r}
install.packages('gridExtra') 
library(gridExtra)

p1 <- ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram()
p2 <- ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram() +
  scale_x_log10()
p3 <- ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram() +
  scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1)
```

### Note instructor's answers with qplot
```{r}
q1 <- qplot(x = friend_count, data = pf)
# Adding +1 helps deal with -Infinity and Infinity as log10 approaches 0
q2 <- qplot(x = log10(friend_count + 1), data = pf)
q3 <- qplot(x = sqrt(friend_count), data = pf)
grid.arrange(q1, q2, q3, ncol=1)

# With the ggplot syntax, I can just use the original plot and add the log10 or sqrt function:
r1 <- ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram()
r2 <- r1 + scale_x_log10()
r3 <- r1 + scale_x_sqrt()
# super convenient & saved me so many lines
grid.arrange(r1, r2, r3, ncol=1)
```

