Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
```

***

#### What are some things that you notice about this histogram?
Response: 0 to 1 has a huge spike compared to the other days. It probably means there are missing data because nobody can be born on day 0.

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: costume party in nz

#### How many of your friends do you think saw that post?
Response: 5

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 1

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
Jan 1st default answer for people who don't submit a birthday
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response: bad data about an extreme case

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response: There is a massive spike in around the first interview, 0 to whatever the next bin is, and a long tail (i.e. skewed right).

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: The max is such a large number it is inflatng the mean

***

### Tenure
Notes:

```{r Tenure}
ggplot(data = pf, aes(x = tenure)) +
    geom_histogram()
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(data = pf, aes(x = tenure/365)) +
    geom_histogram() +
    scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(data = pf, aes(x = tenure/365)) +
    geom_histogram() +
    scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7)) +
    xlab("Number of years using Facebook") +
    ylab("Number of users in sample")

ggplot(aes(x = tenure / 365), data = pf) + 
  geom_histogram(color = 'black', fill = '#F79420') + 
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) + 
  xlab('Number of years using Facebook') + 
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}
ggplot(data = pf, aes(x = age)) +
    geom_histogram()

ggplot(aes(x = age), data = pf) + 
  geom_histogram(binwidth = 1, fill = '#5760AB') + 
  scale_x_discrete(breaks = seq(0, 113, 5))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!