---
title: "ProblemSet3"
author: "Anna Li"
date: "March 15, 2015"
output: html_document
---

# 1. Explore dataset 
```{r}
library(ggplot2)
data(diamonds) # diamonds dataset comes from ggplot2
summary(diamonds)
diamonds$color # prints out the colors for diamonds data set. 
# Levels: D < E < F < G < H < I < J signals the ordering of data
# Learn about diamonds dataset
?diamonds
```

# 2. Price Histogram

```{r, echo=FALSE}
# Create a histogram of the price of
# all the diamonds in the diamond data set.

# TYPE YOUR CODE BELOW THE LINE
# =======================================
names(diamonds)
ggplot(aes(x = price), data = diamonds) +
    geom_histogram()
```

# 3. Price Histogram Summary

```{r}
summary(diamonds$price)
```

# 4. Diamond Counts
```{r}
summary(diamonds$price < 500)
summary(diamonds$price < 250)
summary(diamonds$price >= 15000)
```
1729, 0 , 1656

5. Cheaper Diamonds

```{r}
# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won’t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================================
ggplot(aes(x = price), data = diamonds) +
    geom_histogram(binwidth = 50, color = 'black', fill = '#099DD9') +
    scale_x_continuous(lim = c(0, 2500))
    
```

# 6. Price by Cut Histograms & 7. Price by Cut
```{r}
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================
ggplot(aes(x = price), data = diamonds) +
    geom_histogram(binwidth = 500, color = 'black', fill = '#099DD9') +
    facet_wrap(~cut)
by(diamonds$price, diamonds$cut, summary)

by(diamonds$price, diamonds$cut, max)
```

# 8. Scales and Multiple Histograms
```{r}
# In the two last exercises, we looked at
# the distribution for diamonds by cut.

# Run the code below in R Studio to generate
# the histogram as a reminder.

# ===============================================================

qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = 'free_y')
                                                                               
# ===============================================================

# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

?facet_wrap

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

# If you want a hint, check out the Instructor Notes.

# In ggplot
ggplot(aes(x = price), data = diamonds) +
    geom_histogram(binwidth = 500, color = 'black', fill = '#099DD9') +
    facet_wrap(~cut, scales = 'free_y')
```

# 9. Price per Carat by Cut
```{r}
# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

# Submit your final code when you are ready.

# ENTER YOUR CODE BELOW THIS LINE.
# ===========================================================================

ggplot(aes(x = price/carat), data = diamonds) +
    geom_histogram(color = 'black', fill = '#099DD9') +
    facet_wrap(~cut, scales = 'free_y') +
    scale_x_log10()
```

# 10. Price Box Plots

```{r}
# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won’t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# =================================================================
ggplot(aes(x = cut, y = price), data = diamonds) +
    geom_boxplot()
ggplot(aes(x = clarity, y = price), data = diamonds) +
    geom_boxplot()
ggplot(aes(x = color, y = price), data = diamonds) +
    geom_boxplot()
```

# 11. Interquartile Range - IQR
```{r}
# Price range for middle 50% of diamonds with Color D & J
by(diamonds$price, diamonds$color, summary)
# Example of IQR function - remember to specify variable in subset dataframe 
IQR(subset(diamonds, price <1000)$price)
IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)
```

# 12. Price per Carat Box Plots by Color

```{r}
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

# Go to the discussions to
# share your thoughts and to discover
# what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.

# SUBMIT YOUR CODE BELOW THIS LINE
# ===================================================================
ggplot(aes(x = color, y = price/carat), data = diamonds) +
    geom_boxplot()

```

# 13.  Carat Frequency Polygon 
```{r}
ggplot(aes(x = carat), data = diamonds ) + 
    geom_freqpoly(binwidth = .01) +
    scale_x_continuous(lim = c(0, 2.5))
```

# 14. Data Wrangling with R
Refer to the Resources Used EDA with R doc for details

# 15. Gapminder Data
```{r}
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================
read.csv('sugar_consumption.csv', header = TRUE, row.names = 1)

install.packages('xlsx', dependencies = TRUE) 
library(xlsx)
sugar <- read.xlsx('sugar_consumption.xlsx', sheetName = 'Data')

summary(sugar)
names(sugar)
# rename variable to include country & total
install.packages('reshape', dependencies = TRUE) 
library(reshape)
sugar <- rename(sugar, c(NA. = "Country"))
sugar <- rename(sugar, c(NA..1 = "Total"))
# subsetting by using the index
sugar_val <- sugar[, -1]
sugar_val <- sugar_val[, -45]
# subsetting by using the name of the column
sugar_val <- subset(sugar, select = (names(sugar) != 'Country') & (names(sugar) != 'Total') )
names(sugar_val)

sugar$Total <- rowSums(sugar_val, na.rm = TRUE)
sugar$TotalAvg <- sugar$Total/44
str(sugar)
head(sugar)

sugar_complete <- na.omit(sugar)
str(sugar_complete)
head(sugar_complete)
library(ggplot2)

# Bar chart on total sugar consumption for all countries with data
ggplot(aes(x = Country, y = TotalAvg), data = sugar_complete, fill = Country) +
    geom_bar(stat="identity", color = 'black', fill = '#5760AB') +
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
    xlab("Country") + 
    ylab("Avg. sugar consumption (g/person") +
    ggtitle('Average sugar consumption in grams per person from 1961 to 2004 for 151 countries')

# Bar chart to 
sugar_complete$Country

usa <- subset(sugar_complete, sugar_complete$Country == 'United States')
str(usa)
# Get rid of the 'X' in front of all the years
names(usa) <- substr(names(usa), 2, 5)
library(reshape)
usa <- rename(usa, c(ount = "Country")) # rename columns as appropriate
usa <- rename(usa, c(otal = "Total"))
colnames(usa)[47] <- 'TotalAvg'
names(usa)

# Transpose the columns so they are set up in a dataframe ready for time series chart
subset(usa, select = (names(usa) != 'Country') & (names(usa) != 'Total') & (names(usa) != 'TotalAvg'))
usa <- subset(usa, select = (names(usa) != 'Country') & (names(usa) != 'Total') & (names(usa) != 'TotalAvg'))


head(usa)
tail(usa)
str(usa)
colnames(usa)
rownames(usa)
row.names(usa) <- 'Sugar consumption'
row.names(usa)

usa$Year <- row.names(usa)
?substring
usa1 <- t(usa)
usa[-1]
usa <- usa[-1]
usa <- usa[]
usarows <- rownames(usa)
do.call(rbind, lapply(usa, usa, stringsAsFactors=FALSE)))

usa$Year <- usarows
usadf = as.data.frame(as.table(usa1))
usadf <- rename(usadf, c(Var1 = "Year")) # rename columns as appropriate
usadf <- rename(usadf, c("Sugar Consumption (grams per person)" = "Consumption"))

ggplot(data = usadf, aes(x = Year, y = Consumption, group = Var2)) + 
    geom_line() +
    ggtitle('Sugar Consumption in U.S. from 1961 to 2004') +
    xlab('Year') +
    ylab('Sugar Consumption in U.S. (grams per person)') +
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

str(sugar_val)
sugar_val$Country <- sugar$Country
names(sugar_val) <- substr(names(sugar_val), 2, 5)
sugar_complete <- na.omit(sugar_val)
str(sugar_complete)

# Top 5 countries by total average sugar consumption
sugar_top5 <- subset(sugar, TotalAvg > 150)
# Frequency polygon for sugar data
ggplot(aes(x = Country, y = TotalAvg), data = sugar_top5 ) + 
    geom_bar(stat = "identity") +
    ggtitle("Top 5 countries by average sugar consumption per person from 1961 to 2004") +
    xlab("Country") +
    ylab("Sugar Consumption (g/person)") 

sugar_top5_line <- sugar_top5
names(sugar_top5_line) <- substr(names(sugar_top5), 2, 5)
sugar_top5_line <- rename(sugar_top5_line, c(ount = "Country"))
str(sugar_top5_line)

sugar_top5_line_trans <- t(sugar_top5_line)
sugar_top5_line_trans = as.data.frame(as.table(sugar_top5_line_trans))
sugar_top5_line_trans <- subset(sugar_top5_line_trans, Var1 != 'Country' & Var1 != 'otal' & Var1 != 'otalAvg')
sugar_top5_line_trans$Var2 <- as.factor(sugar_top5_line_trans$Var2)
library(plyr)
sugar_top5_line_trans$Var2 <- revalue(sugar_top5_line_trans$Var2, c("20" = "Barbados"))
sugar_top5_line_trans$Var2 <- revalue(sugar_top5_line_trans$Var2, c("53" = "Costa Rica"))
sugar_top5_line_trans$Var2 <- revalue(sugar_top5_line_trans$Var2, c("56" = "Cuba"))
sugar_top5_line_trans$Var2 <- revalue(sugar_top5_line_trans$Var2, c("101" = "Iceland"))
sugar_top5_line_trans$Var2 <- revalue(sugar_top5_line_trans$Var2, c("241" = "United States"))
sugar_top5_line_trans[sugar_top5_line_trans$Var2 == "20"] <- "Barbados"
sugar_top5_line_trans <- rename(sugar_top5_line_trans, c(Freq = "Consumption"))
sugar_top5_line_trans <- rename(sugar_top5_line_trans, c(Var1 = "Year"))
sugar_top5_line_trans <- rename(sugar_top5_line_trans, c(Var2 = "Country"))
sugar_top5_line_trans$Consumption <- as.numeric(sugar_top5_line_trans$Consumption)

ggplot(data = sugar_top5_line_trans, aes(x = Year, y = Consumption, group = Country)) + 
    geom_line(aes(colour=Country, group=Country)) +
    scale_y_continuous() +
    ggtitle("Average Sugar Consumption per person for Top 5 countries from 1961 to 2004") +
    xlab('Year') +
    ylab('Sugar Consumption (grams per person)') +
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


# See the top 5 countries with highest consumption & lowest consumption, total & avg
# Top 5 country with highest total consumption of sugar
head(arrange(sugar, desc(Total)), 5)
# Top 5 country with highest total consumption of sugar
head(arrange(sugar, desc(TotalAvg)), 5)
## -> U.S., Cuba, Costa Rica, Iceland, Barbados

# Top 5 country with lowest total consumption of sugar
tail(arrange(na.omit(sugar), desc(Total)), 5)
# Top 5 country with lowest total consumption of sugar
tail(arrange(na.omit(sugar), desc(TotalAvg)), 5)
## -> Congo, Dem. Rep., Nepal, Burundi, Laos, Rwanda
```
