Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: Color doesn't always tell us more info
It's OK to reach a dead end in exploratory data analysis

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}

# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

    
# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.

# DO NOT DELETE THESE NEXT TWO LINES OF CODE
# ==============================================================
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
suppressMessages(library(dplyr))

pf.fc_by_age_gender <- pf %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count), 
              median_friend_count = median(as.numeric(friend_count)), n = n())

# ANSWER from instructors: filters out NA and ungroups by gender, then arranges by age.
pf.fc_by_age_gender <- pf %>%
    filter(!is.na(gender)) %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count), 
              median_friend_count = median(as.numeric(friend_count)), n = n()) %>%
    ungroup() %>%
    arrange(age)
pf.fc_by_age_gender

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram(stat="identity")



```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.

ggplot(data = pf.fc_by_age_gender, aes(x = age, y = median_friend_count)) +
    geom_line(aes(color = gender))

head(pf.fc_by_age_gender)
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
install.packages('tidyr', dependencies = TRUE)
library(tidyr)
# dplyr dataframe code creates wide frame with median friend counts
pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>%
  group_by(age) %>%
  arrange(age, gender) %>%
  summarise(female_median = first(median_friend_count),
            male_median = last(median_friend_count))

# wide dataframe correctly created using reshape2 to do same as above
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender, 
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
names(pf.fc_by_age_gender)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.
pf.fc_by_age_gender.wide$ratio <- pf.fc_by_age_gender.wide$female_median/pf.fc_by_age_gender.wide$male_median
head(pf.fc_by_age_gender.wide)

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = ratio)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept = 1, linetype = 2, color = 'red')
    

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

head(pf$tenure, 20)

pf$year_joined <- trunc(abs(pf$tenure/365 - 2014))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

str(pf$year_joined)
?cut

pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.
str(pf)

ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x =  age, y = friend_count)) +
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```


### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x =  age, y = friend_count)) +
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) +
    geom_line(stat = 'summary', fun.y = mean, linetype = 5) 

```

***

### Friending Rate
Notes:

```{r Friending Rate}
# Subset dataframe to only include users with tenure of one day or greater. Then calculate rate

# median friend rate
?median
median(subset(pf$friend_count/pf$tenure, pf$tenure > 0), na.rm = TRUE)

?max
max(subset(pf$friend_count/pf$tenure, pf$tenure > 0), na.rm = TRUE)

```

***

### Friendships Initiated
Notes:

What is the median friend rate?
82

What is the maximum friend rate?
4923

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

ggplot(data = subset(pf, pf$tenure > 0), aes(x =  tenure, y = friendships_initiated/tenure)) +
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)  

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))


```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:
```{r Yogurt dataset}
getwd()
setwd("/Users/horsepower/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson5")

yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)
```

***

### Histograms Revisited

```{r Histograms Revisited}
# more bins so see the gaps between discrete bars & the valleys between tall bars 
ggplot(data = yo, aes(x =  price)) +
    geom_histogram()

# more biased model because fewer number of bars obscures the differences in between 
ggplot(data = yo, aes(x =  price)) +
    geom_histogram(binwidth = 10)

# See these observations numerically instead of graphically
summary(yo)
length(unique(yo$price))
table(yo$price)
```

What do I notice?:
The highest priced yogurts are the most common purchased compared to the cheapest yogurts. There's an unexpected trend around $30-40 that doesn't conform to this trend.
***

### Number of Purchases
Notes:

```{r Number of Purchases}
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!
# OR you can figure out another way to create the
# variable.

?transform
head(airquality,10)
head(transform(airquality, Ozone = -Ozone))
head(transform(airquality, new = -Ozone, Temp = (Temp-32)/1.8))

attach(airquality)
head(transform(Ozone, logOzone = log(Ozone))) # marginally interesting ...
detach(airquality)

head(yo)
names(yo)
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
head(transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry), 10)
head(transform(yo, all.purchases = rowSums(yo[, 4:8])), 10)


```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

