Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes: 

***

### Scatterplots
Notes:

```{r Scatterplots}
getwd()
setwd( "/Users/horsepower/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson3")
pf <- read.delim('pseudo_facebook.tsv') 
library(ggplot2)
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point()
```

***

#### What are some things that you notice right away?
Response: Most of the data is clustered towards age of less than 30 and friend count less than 2000

***

### ggplot Syntax
Notes: The vast majority of people have less than 1000 friends regardless of age. And the majority of users are less than 50 years only, in fact, a huge number are less than 25 years old.

```{r ggplot Syntax}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_jitter(alpha = 1/20) +
    xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20) +
    xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes: The transformation combined with jitter might cause some friend_counts to be negative (which is impossible because you can't have negative friends). So more elaborate syntax is needed to only jitter the ages and not friend_count (see 2nd plot for answer).

```{r Coord_trans()}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20) +
    xlim(13, 90) +
    coord_trans(y = 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0))  +
    xlim(13, 90) +
    coord_trans(y = 'sqrt')
```

#### What do you notice?
There's almost a bimodal distribution of friend count where the under 25s have the first big peak and there's a dip between 30-50 and then another smaller peak around 65.
***

### Alpha and Jitter
Notes: Explore relationship between friends initiated, friendships_initiated (y) and age (x) with ggplot
# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.

```{r Alpha and Jitter}
ggplot(data = pf, aes(x = age, y = friendships_initiated)) +
    geom_point(alpha = 1/5, position = position_jitter(h = 0)) +
    xlim(13, 90)
```

#### What are your observations about your final plot?
The biggest group of friend initiators are in the under 25 group. There is also a sudden spike in the mid to late 60s. The trend is consistent with the friend_count and age plot. 

***

# dplyr
```{r}
install.packages("dplyr", dependencies = TRUE)
library(dplyr)

```
Great library for slicing dataframes and computing quick analyses,  

install.packages("dplyr")

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes: Use the dplyr library. Methods:
group_by()
filter()
mutate()
summarise()
arrange()
Chain them together using %>%
```{r Conditional Means}
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================

pf.fc_by_age <- pf %>%
    group_by(age) %>%
    summarise(friend_count_mean = mean(friend_count),
              n = n()) %>%
    arrange(age)

head(pf.fc_by_age)

```

Create your plot!

```{r Conditional Means Plot}

ggplot(data = pf.fc_by_age, aes(x = age, y = friend_count_mean)) +
    geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'blue')  +
    coord_cartesian(xlim = c(13, 90)) +
    coord_trans(y = 'sqrt') +
    geom_line(stat = 'summary', fun.y = mean, linetype = 1, color = 'red') +
    geom_line(stat = 'summary', fun.y = quantile, probs = .1, linetype = 1, color = 'orange') +
    geom_line(stat = 'summary', fun.y = quantile, probs = .5, linetype = 2, color = 'orange') +
    geom_line(stat = 'summary', fun.y = quantile, probs = .9, linetype = 3, color = 'orange')
    

```

#### What are some of your observations of the plot?
Response:
The mean isn't an accurate description of where most friend count is for age because it is being pull higher by very large outlier friend counts. The quantiles are better measures
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes: According to instructor, 0.3 or -0.3 is meaningful correlation but small. 0.5 or -0.5 is moderate, 0.7 or -0.7 is strong, and 0.9 or -0.9 is large.

```{r Correlation}
cor.test(pf$age, pf$friend_count)

# alternative way to code pearson test with new with() method
with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(filter(pf, age <= 70), cor.test(age, friend_count))
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================

names(pf)
ggplot(data = pf, aes(x = log10(www_likes_received), y = log10(likes_received))) +
    geom_point(alpha = 1/20, color = 'blue') +
    coord_cartesian(xlim = c(0, 6), ylim = c(0,50000)) +
    coord_trans(x = 'sqrt', y = 'sqrt')
    
head(pf$likes_received)

```

***

### Strong Correlations
Notes: Instructor's answer. Much better way of zooming in on biggest cluster of points

```{r Strong Correlations}
ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) +
    geom_point() +
    xlim(0, quantile(pf$www_likes_received, 0.95)) +
    ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response: 0.9479902, rounded to 0.948

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3', dependencies = TRUE)
library(alr3)
data(Mitchell)
summary(Mitchell)
str(Mitchell)
head(Mitchell)
```

Create your plot!

```{r Temp vs Month}
# Create a scatterplot of temperature (Temp)
# vs. months (Month).

ggplot(data = Mitchell, aes(x = Month, y = Temp)) +
    geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0
b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.057
```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
# Break up the x-axis into 12 month increments so it corresponds to a year.
library(ggplot2)
ggplot(data = Mitchell, aes(x = Month %% 12, y = Temp)) +
    geom_point() 
ggplot(data = Mitchell, aes(x = Month, y = Temp)) +
    geom_point() + 
    scale_x_discrete(breaks = seq(0, 203, 12))
install.packages('energy', dependencies = TRUE)
library(energy)
dcor.ttest(Mitchell$Month, Mitchell$Temp, distance = FALSE)
```

***

### A New Perspective

What do you notice?
Response: Within the 12 months, there is an inverse parabolic curve where the highest temperature is in the middle of the year, around 6 month (June/July) mark.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
# Create a new variable, 'age_with_months', in the ‘pf’ data frame.
# Be sure to save the variable in the data frame rather than creating
# a separate, stand-alone variable. You will need to use the variables
# 'age' and 'dob_month' to create the variable ‘age_with_months’.

# Assume the reference date for calculating age is December 31, 2013.
pf <- read.delim("/Users/horsepower/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson3/pseudo_facebook.tsv")
names(pf)
head(pf$dob_month, 10)
head(pf$age, 10)

pf$age_with_months <- pf$age + (1-pf$dob_month/12)

```

***

### Age with Months Means

```{r Age with Months Means}
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data frame should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean    friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11

<!-- 
# Hint 1: Use the group_by(), summarise(), and arrange() functions in the dplyr package to split the data frame by age_with_month. Make sure you arrange by the correct variable (it's not age anymore). 

# Hint 2: The code should look similar to the code when we split the data frame by age and found summaries. age_groups <- group_by(pf, age) 
# pf.fc_by_age <- summarise(age_groups, 
#                                                     friend_count_mean = mean(friend_count), 
#                                                     friend_count_median = median(friend_count),
#                                                     n = n()) 
# pf.fc_by_age <- arrange(pf.fc_by_age, age) 
# 
# head(pf.fc_by_age)
-->
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_months <- pf %>%
    group_by(age_with_months) %>%
    summarise(friend_count_mean = mean(friend_count), 
              friend_count_median = median(friend_count),
              n = n()) %>%
    arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

