Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes: 

***

### Scatterplots
Notes:

```{r Scatterplots}
getwd()
setwd( "/Users/horsepower/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson3")
pf <- read.delim('pseudo_facebook.tsv') 
library(ggplot2)
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point()
```

***

#### What are some things that you notice right away?
Response: Most of the data is clustered towards age of less than 30 and friend count less than 2000

***

### ggplot Syntax
Notes: The vast majority of people have less than 1000 friends regardless of age. And the majority of users are less than 50 years only, in fact, a huge number are less than 25 years old.

```{r ggplot Syntax}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_jitter(alpha = 1/20) +
    xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20) +
    xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes: The transformation combined with jitter might cause some friend_counts to be negative (which is impossible because you can't have negative friends). So more elaborate syntax is needed to only jitter the ages and not friend_count (see 2nd plot for answer).

```{r Coord_trans()}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20) +
    xlim(13, 90) +
    coord_trans(y = 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(data = pf, aes(x = age, y = friend_count)) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0))  +
    xlim(13, 90) +
    coord_trans(y = 'sqrt')
```

#### What do you notice?
There's almost a bimodal distribution of friend count where the under 25s have the first big peak and there's a dip between 30-50 and then another smaller peak around 65.
***

### Alpha and Jitter
Notes: Explore relationship between friends initiated, friendships_initiated (y) and age (x) with ggplot
# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.

```{r Alpha and Jitter}
ggplot(data = pf, aes(x = age, y = friendships_initiated)) +
    geom_point(alpha = 1/5, position = position_jitter(h = 0)) +
    xlim(13, 90)
```

#### What are your observations about your final plot?
The biggest group of friend initiators are in the under 25 group. There is also a sudden spike in the mid to late 60s. The trend is consistent with the friend_count and age plot. 

***

# dplyr
```{r}
install.packages("dplyr", dependencies = TRUE)
library(dplyr)

```
Great library for slicing dataframes and computing quick analyses,  

install.packages("dplyr")

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

