---
title: "ResourcesUsedEDAwithR"
author: "Anna Li"
date: "March 11, 2015"
output: html_document
---

# Common & Useful R commands 
## syntax for creating R chunks in RMD
```{r}
summary(mtcars)
# Find the 90% quartile of data in variable prices in diamonds dataset
quantile(diamonds$price, 0.9)
```

### Remove command output from knitted HTML file
http://stackoverflow.com/questions/13090838/r-markdown-avoiding-package-loading-messages
```{r include=FALSE, message=FALSE, cache=FALSE}
```

## Comment blocks
Command + Shift + c (but only for the r code blocks)
http://stackoverflow.com/questions/4131338/is-it-possible-to-have-a-multi-line-comments-in-r

## Directories
### Find current directory
getwd()
### Move to a different directory
setwd('/Users/horsepower/Dropbox/Udacity/NanoDA/EDAwithR/EDA_Course_Materials/lesson2')
### See all files in directory
list.files()

## CSV
reddit <- read.csv('reddit.csv')
*If the file is tab separated, there are two ways to import the data:*   
reddit <- read.csv('reddit.tsv', sep='\t')
reddit <- read.delim('reddit.tsv')
*Additional arguments to determine header & check variable names*
reddit <- read.delim('reddit.tsv', row.names = 1, check.names = F)


## Exploring data
**Column names (in R times, see the variables)**
names(reddit)
**Find the dimensions (rows, columns)**
dim(reddit)
**Find the structure (columns, data types, example of values...)**
str(reddit)
**Print out the first 6 lines of dataframe (6 is default)**
head(reddit)
head(reddit, 3) # integer specifies only first 3 lines, use negative to last 3 lines
**Print out last 6 lines of dataframe**
tail(reddit)
tail(reddit, 10) # print out last 10 lines, can use negative integer
**Descriptive stats on data, e.g. mean, median, & quartiles **
summary(reddit)
**Find all attributes**
attributes(reddit)
**Find all column names**
colnames(reddit)
**Find all row names**
rownames(reddit)
**See all possible values for a factor**
levels(reddit)
**See all unique values in a dataframe, vector, array, or column**
unique(reddit)
https://stat.ethz.ch/R-manual/R-devel/library/base/html/unique.html



## Basic Plotting
### qplot
qplot(x = friend_count, data = pf)
(x is the x variable you want to display in the histogram)
### ggplot 
ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram()
  
## ggplot extras
*Increase size of font on axes*
axis_sizer <- theme(axis.title.x = element_text(size = 18),
                    axis.title.y = element_text(size = 18),
                    axis.text.x = element_text(size = 16),
                    axis.text.y = element_text(size = 16))
                    
## dplyr how tos
*Sample smaller subset of data for quicker processing*
diamonds <- sample_n(diamonds, 10000)
# Resources Used in the Exploratory Data Analysis with R class

## Installation of packages
* Error around mirroring
http://stackoverflow.com/questions/16826933/installation-of-r-fpc-package
http://stackoverflow.com/questions/17705133/package-error-when-running-r-code-on-command-line

## ggplot
### Facet_wrap
http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
### Box plot
http://stackoverflow.com/questions/21533158/remove-outliers-fully-from-multiple-boxplots-made-with-ggplot2-in-r-and-display
http://docs.ggplot2.org/0.9.3.1/geom_boxplot.html
### Frequency Polygon
http://www.inside-r.org/packages/cran/ggplot2/docs/geom_freqpoly
http://stackoverflow.com/questions/21121446/how-to-plot-in-the-same-graph-the-histogram-and-the-frequency-polygon-of-two-s
### Histograms
Free scales
http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
### Line plots
http://docs.ggplot2.org/current/geom_line.html
http://stackoverflow.com/questions/10357768/plotting-lines-and-the-group-aesthetic-in-ggplot2
### Scale x axis
**log10**
http://forums.udacity.com/questions/100224132/error-using-scale_x_log10-with-this-data
### Scale y axis
**scale_y_continuous**
http://www.cookbook-r.com/Graphs/Axes_(ggplot2)/
### Setting labels
http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_(ggplot2)/
http://www.cookbook-r.com/Graphs/Titles_(ggplot2)/
http://stackoverflow.com/questions/10438752/adding-x-and-y-axis-labels-in-ggplot2
### Styling
**Add color to lines** 
http://www.cookbook-r.com/Graphs/Colors_%28ggplot2%29/
**Add color to bars in histogram**
http://stackoverflow.com/questions/17732084/colours-for-geom-histogram
http://docs.ggplot2.org/current/geom_histogram.html
http://stackoverflow.com/questions/10016316/how-to-manually-fill-colors-in-a-ggplot2-histogram
**Add color to points in scatterplot**
http://docs.ggplot2.org/0.9.3.1/geom_point.html
### Plotting Datetime
http://matpalm.com/blog/2012/03/18/ggplot_posix_cheat_sheet/
### Coord_cartesian & Coord_trans
http://docs.ggplot2.org/current/coord_trans.html
### Stat Summary
http://docs.ggplot2.org/current/stat_summary.html
### Aesthetics
http://docs.ggplot2.org/current/aes_linetype_size_shape.html
### Smoothers
http://www.ats.ucla.edu/stat/r/faq/smooths.htm
### GGally 
**Scatterplot Matrix (ggpairs)**
http://www.gettinggeneticsdone.com/2011/07/scatterplot-matrices-in-r.html
### Add Horizontal or Vertical line to plot
http://www.sthda.com/english/wiki/ggplot2-add-straight-lines-to-a-plot-horizontal-vertical-and-regression-lines
### Rotate axis labels
http://stackoverflow.com/questions/10063701/how-to-rotate-the-axis-labels-in-ggplot2
http://stackoverflow.com/questions/15838533/rotating-axis-labels-in-date-format

## R language syntax
### Calculate percentage statistics on a column in a dataframe
http://stackoverflow.com/questions/9623763/in-r-how-can-i-compute-percentage-statistics-on-a-column-in-a-dataframe-tabl
### Read in dataset into R
http://stackoverflow.com/questions/18161009/error-in-reading-in-data-set-in-r
http://www.statmethods.net/input/importingdata.html
### More about Factors in R
http://www.stat.berkeley.edu/~s133/factors.html
#### rowSums
http://stackoverflow.com/questions/3991905/vector-that-is-sum-of-rows
http://stackoverflow.com/questions/20819100/colsum-rowsum-populating-matrix
http://www.inside-r.org/r-doc/base/colMeans
### Rename variables
http://www.statmethods.net/management/variables.html
http://stackoverflow.com/questions/7531868/how-to-rename-a-single-column-in-a-data-frame-in-r
### Substring
http://stackoverflow.com/questions/7963898/extracting-the-last-n-characters-from-a-string-in-r
### Create a dataframe from a matrix
http://stackoverflow.com/questions/15885111/create-data-frame-from-a-matrix-in-r
### Subsetting
http://stackoverflow.com/questions/8005154/conditionally-remove-dataframe-rows-with-r
### Overwrite/Replace values
http://stackoverflow.com/questions/19503266/r-replace-all-particular-values-in-a-data-frame
http://stackoverflow.com/questions/5824173/replace-a-value-in-a-data-frame-based-on-a-conditional-if-statement-in-r
### plyr library
**subsetting & deleting rows**
http://blog.rstudio.org/2014/01/17/introducing-dplyr/
**arrange**
http://127.0.0.1:24449/library/plyr/html/arrange.html
http://stackoverflow.com/questions/10879551/frequency-count-of-two-column-in-r
### dplyr
http://blog.rstudio.org/2014/01/17/introducing-dplyr/
http://rstudio-pubs-static.s3.amazonaws.com/11068_8bc42d6df61341b2bed45e9a9a3bf9f4.html
http://datascience.la/hadley-wickhams-dplyr-tutorial-at-user-2014-part-1/
http://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-2/
#### code examples
### lapply
https://rollingyours.wordpress.com/2014/10/20/the-lapply-command-101/
#### First intro to list
https://rollingyours.wordpress.com/2014/10/20/the-lapply-command-101/
***
### Quantile
https://stat.ethz.ch/R-manual/R-devel/library/stats/html/quantile.html

http://rstudio-pubs-static.s3.amazonaws.com/11068_8bc42d6df61341b2bed45e9a9a3bf9f4.html
### Date Time methods 
**strptime**
http://127.0.0.1:24449/library/base/html/strptime.html
http://stackoverflow.com/questions/9372211/reading-csv-with-date-and-time
**lubridate**
http://cran.r-project.org/web/packages/lubridate/lubridate.pdf

### Math
** Difference between floor, ceiling & truncate **
http://stackoverflow.com/questions/14/difference-between-math-floor-and-math-truncate
** Built in functions **
http://www.statmethods.net/management/functions.html
### Cut
http://www.r-bloggers.com/r-function-of-the-day-cut-2/

# Data Wrangling Resources
Taken directly from Udacity's EDA with R course materials
https://www.udacity.com/course/viewer#!/c-ud651-nd/l-685569242/m-3552939568
> Data munging or data wrangling can take up much of a data scientist's or data analyst's time. There are 
> two R packages that make these tasks easier in R: tidyr and dplyr.

> tidyr -a package that reshapes the layout of your data

> dplyr - a package that helps you transform tidy, tabular data

> Review Data Wrangling in R to get a sense of how these packages allow you to manipulate data. You can use > these packages to help you in the next programming task and in your future investigations.
> http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
> http://www.rstudio.com/resources/cheatsheets/
> You may also download this useful Data Wrangling Cheat Sheet. There are some other useful cheat sheets at > RStudio. https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf

## From wide to long and long to wide
### tidyr & reshape2
http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/
### reshape2
http://seananderson.ca/2013/10/19/reshape.html
http://www.r-bloggers.com/melt/

## Functions
### 3 different kinds of functions & their syntax
http://discussions.udacity.com/t/writing-functions-in-r/14299/3


## Linear Regression in R
http://data.princeton.edu/R/linearModels.html
### Interpreting regression coefficients in R
http://www.r-bloggers.com/interpreting-regression-coefficient-in-r/?utm_source=feedburner&utm_medium=email&utm_campaign=Feed%3A+RBloggers+%28R+bloggers%29
http://www.theanalysisfactor.com/interpreting-regression-coefficients/
http://blog.yhathq.com/posts/r-lm-summary.html
http://stats.stackexchange.com/questions/24242/how-to-apply-coefficient-term-for-factors-and-interactive-terms-in-a-linear-equa/24256#24256

# Data Wrangling Help
## Pandas Data Frame wrangling
http://stackoverflow.com/questions/10665889/how-to-take-column-slices-of-dataframe-in-pandas
### Write pandas data frame to csv
http://stackoverflow.com/questions/16923281/pandas-writing-dataframe-to-csv-file
### Perform a join with two dataframes
https://stat.ethz.ch/R-manual/R-devel/library/base/html/merge.html
http://stackoverflow.com/questions/1299871/how-to-join-data-frames-in-r-inner-outer-left-right
